---
title: "Code (Life Expectancy)"
author: "Group 6"
date: "2022-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load Package

```{r}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(corrplot)
library(leaps)
library(car)
library(Metrics)
library(reshape2)
library(ggpubr)
library(moments)
library(naniar)
library(VIM)
library(plotly)
```

## Part 1. Load data

```{r}
data_input<-read_csv("./data/Life Expectancy Data.csv")
str(data_input)
```

```{r}
head(data_input)
```

## Part 2. Clean and filter data

### 2.1 Missing data

```{r}
missing.rows = dim(data_input)[1] -  dim(na.omit(data_input))[1]

missings_df <- data.frame(type=c("missing", "non-missing") ,count = c(missing.rows,  dim(na.omit(data_input))[1]))

ggplot(missings_df, aes(fill=type, y="", x=count)) +
    geom_bar(position="stack", stat="identity")+
    ggtitle("Missing vs Non-missing row counts") +
    xlab("Missing count") + ylab("") 
```

```{r}
vis_miss(data_input)
```

```{r}
data_fill <- kNN(data_input)
data_fill <- subset(data_fill, select = country:schooling)
vis_miss(data_fill)
```

### 2.2 **Remove unnecessary variables and** **factorize categorical variables**:

```{r}
data<-data_fill[-c(1,2,3)]
glimpse(data)
```

### **2.3 Data transformation**

```{r}
ggplot(data, aes(x=life_expectancy)) + 
    geom_density(alpha=.3, fill="red", color="red", size=1.5)+
    geom_vline(aes(xintercept=mean(life_expectancy)), size=1)+
    ggtitle("Distribution density of Life.expectancy") +
    theme(text = element_text(size = 18))
```

```{r}
sprintf("Skewness: [%s]", toString(skewness(data$life_expectancy, na.rm = TRUE)))
```

```{r}
data$life_expectancy <- sqrt(max(data$life_expectancy+1)- data$life_expectancy)

ggplot(data, aes(x=life_expectancy)) +
    geom_density(alpha=.3, fill="red", color="red", size=1.5)+
    geom_vline(aes(xintercept=mean(life_expectancy)))+
    ggtitle("Distribution density of Life.expectancy") +
    theme(text = element_text(size = 18))
```

## Part 3. **Exploratory Data Analysis (EDA)**

### **3.1 Correlation and collinearity**

```{r}
corr <- round(cor(data), 3)
ggcorrplot(corr,type = "upper", lab = TRUE, , outline.color = "black", lab_size = 1, legend.title = "Correlation")+
ggtitle("Correlation Matrix")
```

```{r}
model_linear <- lm(life_expectancy~ ., data = data)
vifs <- data.frame(vif(model_linear))

ggplot(vifs, aes(y=vif.model_linear., x=row.names(vifs))) +
    geom_bar(aes(fill=vif.model_linear.>5),stat="identity")+
    scale_y_continuous(trans = "sqrt",  breaks = c(5, 10, 50, 100))+
    geom_hline(yintercept = 5, colour = "red") +
    ggtitle("VIF per feature") +
    xlab("Featurs") + ylab("VIF") +
    theme(axis.text.x=element_text(angle=40, hjust=1))
```

```{r}
data_EDA <- subset(data, select = -c(infant_mortality, `thinness_5-9`
, GDP))

model_EDA <- lm(life_expectancy~ ., data = data_EDA)
vifs_EDA <- data.frame(vif(model_EDA))


ggplot(vifs_EDA, aes(y=vif.model_EDA., x=row.names(vifs_EDA))) +
    geom_bar(aes(fill=vif.model_EDA.>5),stat="identity")+
    scale_y_continuous(trans = "sqrt",  breaks = c(5, 10, 50, 100))+
    geom_hline(yintercept = 5, colour = "red") +
    ggtitle("VIF per feature") +
    xlab("Featurs") + ylab("VIF") +
    theme(axis.text.x=element_text(angle=40, hjust=1))
```

## Part 4. Data visualization

```{r}
data_visu <- read_csv("./data/data_Visu.csv")
data_visu <- na.omit(data_visu)
```

```{r}
g <- list(
showframe = TRUE,
showcoastlines = TRUE,
projection = list(type = 'Mercator')
)

plot <- plot_geo(data_fill, locationmode="country names")%>%
    add_trace(locations=~country,
             z=~life_expectancy,
             frame=~year) %>%
    layout(
  title = with(data, paste('Mean Life Expectancy')),
  geo = g
)
plot
```

```{r}
fig1<-plot_ly(data_visu,x=~schooling,y=~life_expectancy,size=~population,frame = ~year,text=~country,color = ~Continent,log_x=TRUE,size_max=60)
fig1
```

```{r}
fig2<-plot_ly(data_visu,x=~percent_expenditure,y=~life_expectancy,size=~population,frame = ~year,text=~country,color = ~Continent,log_x=TRUE,size_max=60)
fig2
```

```{r}
ggplot(data_visu,aes(x=GDP,y=life_expectancy))+geom_point(color="red")+geom_smooth(method="lm")+facet_wrap(~Continent)
```

```{r}
ggplot(data_visu,aes(x=income_composition_of_resources,y=life_expectancy))+geom_point(color="blue")+geom_smooth(method="lm")+facet_wrap(~Continent)
```

## Part 5. Models

## 1. Model (full)

Sử dụng data đầy đủ để tạo model với tất cả cột

```{r}
set.seed(20133076)
sample_full<-sample(c(TRUE,FALSE),nrow(data),replace=TRUE, prob = c(0.7,0.3))

train_full<-data[sample_full,]
x_test_full<-data[!sample_full,]
y_test_full<-data[!sample_full,]$life_expectancy

model_full<-lm(life_expectancy~.,data=train_full)

summary(model_full)
```

Kiểm tra

```{r}
pred_full<-predict(model_full, newdata = x_test_full)
value_model <- data.frame("Full", rmse(pred_full,y_test_full), summary(model_full)$adj.r.squared,length(coef(model_full))-1)

colnames(value_model) = c("Model", "Root Mean Squared Error", "R squared", "Number of variable")

value_model
```

```{r}
par(mfrow=c(2,2))
plot(model_full)
```

```{r}
pred_acct_full <- data.frame(pred_full, y_test_full)
ggplot(pred_acct_full,aes(x=pred_full, y= y_test_full)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
```

## 2. Model (AIC)

```{r}
set.seed(20133082)
sample_step<-sample(c(TRUE,FALSE),nrow(data),replace=TRUE, prob = c(0.7,0.3))

train_step<-data[sample_step,]
x_test_step<-data[!sample_step,]
y_test_step<-data[!sample_step,]$life_expectancy

model_full_step<-lm(life_expectancy~.,data=train_step)
```

```{r, results=FALSE}
model_step<-step(model_full_step,direction = "both",test="F")

```

```{r}
summary(model_step)
```

Kiểm tra

```{r}
pred_step<-predict(model_step, newdata = x_test_step)
value_model[nrow(value_model) + 1,] = c("AIC",rmse(y_test_step,pred_step), summary(model_step)$adj.r.squared, length(coef(model_step))-1)
value_model
```

```{r}
par(mfrow=c(2,2))
plot(model_step)
```

```{r}
pred_acct_step <- data.frame(pred_step, y_test_step)
ggplot(pred_acct_step,aes(x=pred_step, y= y_test_step)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
```

## 3. Feature selection methods

Lựa chọn theo 3 phương pháp:

-   Best subset

-   Forward inclusion

-   Backward elimination

theo đó dựa trên:

-   Residual sum of squares (RSS)

-   Adjusted R\^2 (Adjr2)

-   Mallow's Cp (CP)

-   Bayesian information criterion (BIC)

#### 3.1 Best subset

```{r}
set.seed(20133035)
regfit.best <- regsubsets(life_expectancy~., data= data_EDA, nvmax = 18)
reg.summary <- summary(regfit.best)

par(mfrow=c(2,2))

#- residual sum of squares:
plot(reg.summary$rss,xlab="Number of Variables",ylab="RSS",type="l")
which.min(reg.summary$rss)
points(15,reg.summary$rss[15], col="red",cex=2,pch=20)

# adjusted-R^2 with its largest value
plot(reg.summary$adjr2,xlab="Number of Variables",ylab="Adjusted Rsq",type="l")
which.max(reg.summary$adjr2)
points(14,reg.summary$adjr2[14], col="red",cex=2,pch=20)

# Mallow's Cp with its smallest value
plot(reg.summary$cp,xlab="Number of Variables",ylab="Cp",type='l')
which.min(reg.summary$cp)
points(14,reg.summary$cp[14],col="red",cex=2,pch=20)

# BIC with its smallest value
plot(reg.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(reg.summary$bic)
points(10,reg.summary$bic[15],col="red",cex=2,pch=20)
```

Có thể thấy BIC cho số biến ít nhất nên từ đây chỉ xét theo BIC.

#### 3.2 Forward inclusion

```{r}

par(mfrow=c(1,1))
set.seed(20133013)
regfit.fwd <- regsubsets(life_expectancy~.,data=data_EDA,nvmax=18,method="forward")
fwd.summary <-summary(regfit.fwd)

# fwd.summary
#set_plot_dimensions(8,6)
plot(fwd.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(fwd.summary$bic)
points(10,fwd.summary$bic[10],col="red",cex=2,pch=20)
```

#### 3.3 Backward elimination

```{r}
set.seed(20133013)
regfit.bwd <- regsubsets(life_expectancy~.,data=data_EDA,nvmax=18,method="backward")
bwd.summary <- summary(regfit.bwd)

# fwd.summary
#set_plot_dimensions(8,6)
plot(bwd.summary$bic,xlab="Number of Variables",ylab="BIC",type='l')
which.min(bwd.summary$bic)
points(10,bwd.summary$bic[10],col="red",cex=2,pch=20)
```

Tất cả các phương pháp với BIC đều cho ra kết quả là 10 biến.

```{r}
v_names <- rownames(as.data.frame(coef(regfit.best,10)))
coefs<- data.frame(v_names)
coefs$best_coef_value<- coef(regfit.best,10)
coefs$fwd_coef_value <-  coef(regfit.fwd,10)
coefs$bwd_coef_value <-  coef(regfit.bwd,10)

#set_plot_dimensions(18,4)
ggplot(coefs,
       aes(x=v_names, y=best_coef_value, fill=best_coef_value)) +
                                  geom_bar(stat="identity") +
                                  ggtitle("Features & coeffecients: [method Best]") +
                                  xlab("Feature") + ylab("Coef value") +
                                  theme(axis.text.x=element_text(angle=45, hjust=1))

ggplot(coefs,
       aes(x=v_names, y=fwd_coef_value, fill=fwd_coef_value)) +
                                  geom_bar(stat="identity") +
                                  ggtitle("Features & coeffecients: [method Forward inclusion]") +
                                  xlab("Feature") + ylab("Coef value") +
                                  theme(axis.text.x=element_text(angle=45, hjust=1))
ggplot(coefs,
       aes(x=v_names, y=bwd_coef_value, fill=bwd_coef_value)) +
                                  geom_bar(stat="identity") +
                                  ggtitle("Feature & coeffecients: [method Backward elimination]") +
                                  xlab("Feature") + ylab("Coef value") +
                                  theme(axis.text.x=element_text(angle=45, hjust=1))
```

3 phương pháp đã chọn ra 10 biến. Tạo data:

```{r}

data_select<-subset(data, select = c(life_expectancy, adult_mortality, percent_expenditure, BMI, polio, total_expenditure, diphtheria, `HIV/AIDS_mortality`, `thinness_1-19`,income_composition_of_resources,  schooling ))

```

#### 3.4 Final model

```{r}
set.seed(20133077)
sample_select<-sample(c(TRUE, FALSE), nrow(data_select), replace=TRUE, prob=c(0.70,0.30))

train_select<-data_select[sample_select,]
x_select_test<-data_select[!sample_select,]
y_select_test<-data_select[!sample_select,]$life_expectancy


model_select<-lm(life_expectancy~.,data=train_select)
summary(model_select)

```

Kiểm tra

```{r}
pred_select<-predict(model_select, newdata = x_select_test)

value_model[nrow(value_model) + 1,] = c("BIC",rmse(y_select_test,pred_select), summary(model_select)$adj.r.squared, length(coef(model_select))-1)
value_model
```

```{r}
par(mfrow=c(2,2))
plot(model_select)
```

```{r}
pred_acct_select <- data.frame(pred_select, y_select_test)
ggplot(pred_acct_select,aes(x=pred_select, y= y_select_test)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
```
